String Constraints with Regex-Counting and String-Length Solved Efficiently

### 2023-05-13 2023-1-19/sections/implementation.tex
references

\begin{table}[ht]
\vspace{-2mm}
\begin{center}
  \import{tables}{table_regcol.tex}
\end{center}
  \caption{Experiment results on RegCol, with timeout = 60 seconds}
  \label{tab:results_regcol}
\vspace{-5mm}
\end{table}

@@

\begin{table}[ht]
\vspace{-2mm}
\begin{center}
  \import{tables}{table_automatark.tex}
\end{center}
  \caption{Experiment results on AutomatArk, with timeout = 60 seconds}
  \label{tab:results_automatark}
\vspace{-5mm}
\end{table}

@@

In summary, from the results in Table~\ref{tab:results_regcol} and Table~\ref{tab:results_automatark}, $\ostrichrecl$ solves 48,698 out of 48,834 instances, which is 1,091 more than Z3str3RE, the state-of-the-art best solver on RECL constraints. Moreover, the performance (speed) of $\ostrichrecl$ is comparable to that of Z3str3RE.

>>>>

\begin{table}[ht]
\vspace{-3mm}
\begin{center}
  \import{tables}{table_regcol.tex}
\end{center}
  \caption{Experiment results on RegCol, with timeout = 60 seconds}
  \label{tab:results_regcol}
\vspace{-6mm}
\end{table}

@@

\begin{table}[ht]
%\vspace{-1mm}
\begin{center}
  \import{tables}{table_automatark.tex}
\end{center}
  \caption{Experiment results on AutomatArk, with timeout = 60 seconds}
  \label{tab:results_automatark}
\vspace{-6mm}
\end{table}

@@

In summary, $\ostrichrecl$ solves 48,698 out of 48,834 instances, which is 1,091 more than Z3str3RE, the state-of-the-art best solver on RECL constraints. Moreover, the speed of $\ostrichrecl$ is the second fastest and comparable to that of Z3str3RE. (See Table~\ref{tab:results_summary}.)

\begin{table}[ht]
\vspace{-3mm}
\begin{center}
  \import{tables}{table_summary.tex}
\end{center}
  \caption{Experiment results: A Summary}
  \label{tab:results_summary}
\vspace{-6mm}
\end{table}

### 2023-05-13 2023-1-19/sections/introduction.tex
references

The representation of regex-counting and string-length as registers in CEFA, combined with the aforementioned size-reduction and under-approximation techniques, entails an efficient algorithm for solving RECL constraints. We implement the algorithm on top of OSTRICH, resulting in a new version of OSTRICH, called $\ostrichrecl$ for convenience. 
%
Furthermore, we utilize a collection of benchmark suites, which comprise 48,834 RECL constraints in total, to evaluate the performance of $\ostrichrecl$. The collection of benchmark suites includes a widely-used benchmark suite for string constraints, as well as benchmark suites comprising constraints generated from regular expressions in various open source regex libraries. 
%
%Note that in the benchmark suites, we only consider the regexes that contain explicit counting, that is, regexes that contain sub-expressions of the form $e^{\{n, m\}}$, and ignore those that contain only implicit counting operators, that is, $e^*$ or $e^+$, since the state-of-the-art string solvers suffer mainly from the explicit counting operators. 
%
The experiment results show that $\ostrichrecl$ can solve 48,698 out of 48,834 instances, while Z3str3RE, the state-of-the-art best solver on RECL constraints, solves only 47,607 instances (where the timeout period is set as 60 seconds). Moreover, the speed of $\ostrichrecl$ is comparable to that of Z3str3RE. In comparison, CVC5, Z3seq, Z3-Trau, and Z3str3 solve only 44,704, 46,426, 24,121, and 35,866 instances respectively. (See Table~\ref{tab:results_regcol} and Table~\ref{tab:results_automatark} for more details.)

At last, we would like to clarify the relationship of this work with~\cite{atva2020}, where CEFA was introduced. This work differs from \cite{atva2020} in the following aspects: In \cite{atva2020}, regex-counting was unfolded explicitly. Moreover, in~\cite{atva2020}, the nonemptiness of CEFA w.r.t. LIA formulas was solved by utilizing the nuXmv model checker~\cite{nuxmv}. In contrast, this work solves it by integrating the techniques of Parikh-image computation, automata-size reduction, and under approximation. The experimental results show that our approach solves 7,092 more instances than the solver in~\cite{atva2020} while spending less time. (See Table~\ref{tab:results_regcol} and Table~\ref{tab:results_automatark} for more details, where OSTRICH denotes the solver in \cite{atva2020}.)

>>>>

The representation of regex-counting and string-length as registers in CEFA, combined with the aforementioned size-reduction and under-approximation techniques, entails an efficient algorithm for solving RECL constraints. We implement the algorithm on top of OSTRICH, resulting in a new version of OSTRICH, called $\ostrichrecl$ for convenience. 
%
Furthermore, we utilize a collection of benchmark suites, which comprise 48,834 RECL constraints in total, to evaluate the performance of $\ostrichrecl$. The collection of benchmark suites includes a widely-used benchmark suite for string constraints, as well as benchmark suites comprising constraints generated from regular expressions in various open source regex libraries. 
%
%Note that in the benchmark suites, we only consider the regexes that contain explicit counting, that is, regexes that contain sub-expressions of the form $e^{\{n, m\}}$, and ignore those that contain only implicit counting operators, that is, $e^*$ or $e^+$, since the state-of-the-art string solvers suffer mainly from the explicit counting operators. 
%
The experiment results show that $\ostrichrecl$ can solve 48,698 out of 48,834 instances, while Z3str3RE, the state-of-the-art best solver on RECL constraints, solves only 47,607 instances (where the timeout period is set as 60 seconds). Moreover, the speed of $\ostrichrecl$ is comparable to that of Z3str3RE. In comparison, CVC5, Z3seq, Z3-Trau, and Z3str3 solve only 44,704, 46,426, 24,121, and 35,866 instances respectively. (See Table~\ref{tab:results_regcol}-\ref{tab:results_summary} for more details.)

At last, we would like to clarify the relationship of this work with~\cite{atva2020}, where CEFA was introduced. This work differs from \cite{atva2020} in the following aspects: In \cite{atva2020}, regex-counting was unfolded explicitly. Moreover, in~\cite{atva2020}, the nonemptiness of CEFA w.r.t. LIA formulas was solved by utilizing the nuXmv model checker~\cite{nuxmv}. In contrast, this work solves it by integrating the techniques of Parikh-image computation, automata-size reduction, and under approximation. The experimental results show that our approach solves 7,092 more instances than the solver in~\cite{atva2020} while spending less time. (See Table~\ref{tab:results_regcol}-\ref{tab:results_summary} for more details, where OSTRICH denotes the solver in \cite{atva2020}.)
